---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(groebner)
```


## Example 1

Consider the system of polynomials
$$
\begin{gather*}
x^2 - 4 x y + 4 y^2 - 4 \\
4 x^2 - 4 x y + y^2 - 4
\end{gather*}
$$

The `parse_polys` function reads a string representation of a set of polynomials in an arbitrary set
of variables. however, once read the polynomials are always printed in terms of the standard 
variables `x1`, `x2`, `x3`, ....
```{r}
ps <- parse_polys(c("x^2-4*x*y+4*y^2-4",
                    "4*x^2-4*x*y+y^2-4"),
                  c("x","y"))
ps
```

The `groebner` function computes the Groebner basis of the system of polynomials.
```{r}
gb <- groebner(ps)
gb
```
In this case the initial system is already a reduced Groebner basis.

The roots of the system are found with the `solve_polys`
```{r}
solve_polys(ps,gb)
```


### Detail

The `solve_polys` function is a wrapper that derives a monomial basis, and then computes the roots
of the system as the eigenvalues of the matrices that represent multiplication by the standard 
variables in the monomial basis.

The monomial basis is computed by `monomial_basis`.  The basis are represented as a list of integer
vectors specifying the powers of the standard variables in the monomial.
```{r}
ms <- monomial_basis(gb)
ms
```

The matrices that represent multiplication by the standard variables in terms of the monomial basis
are computed by `multiplication_matrix`.
```{r}
Ms <- lapply(1:2,function(k) multiplication_matrix(k,ms,gb))
Ms
```

These matrices commute
```{r}
Ms[[1]]%*%Ms[[2]] - Ms[[2]]%*%Ms[[1]]
```
and a common eigenbasis that will simultaneously diagonalize these matrices is found with 
`common_eigenbasis`.
```{r}
Bs <- common_eigenbasis(Ms)
Bs
```

The candidate roots of the system are given by the eigenvalues corresponding to the vectors in the
eigenbasis
```{r}
rs <- roots(Ms,Bs)
rs
```

The total absolute error at the roots is
```{r}
rowSums(abs(eval_polys(ps,rs)))
```

The roots can be refined with a small number of Newton iterations.
```{r}
rs <- newton_polish(ps,rs,3)
rowSums(abs(eval_polys(ps,rs)))
```

Alternately, if a common left eigenbasis is calculated from the multiplication matrices
```{r}
lBs <- common_eigenbasis(Ms,left=TRUE)
```
then the basis elements are proportional to the values of the monomial basis at the roots, 
and roots can be found directly from the (left) eigenbasis provided that each standard variable
appears as an element of the monomial basis
```{r}
roots_left_eigenbasis(ms,lBs)
```


### Term Orders

By default, computations are performed with respect to a graded reverse lexicographic term order,
but three standard term order functions are available:

* `lex` lexicographic order
* `grlex` graded lexicographic order
* `grevlex` graded reverse lexicographic order

and the term order can be set with `set_term_order`.  But if the term order is changed, any existing
polynomials are left in an inconsistent state and must be redefined with `poly` to obtain valid
polynomials.

With a lexicographic term order,
```{r}
set_term_order(lex)
ps <- lapply(ps,poly)
ps
```

the Groebner basis is
```{r}
gb <- groebner(ps)
gb
```


## Example 2

Consider the system of polynomials
$$
\begin{gather*}
x^2 + y + z -1 \\
x + y^2 + z -1 \\
x + y + z^2 -1 
\end{gather*}
$$

Define the polynomials
```{r}
set_term_order(grevlex)
ps <- parse_polys(c("x^2+y+z-1",
                    "x+y^2+z-1",
                    "x+y+z^2-1"),
                  c("x","y","z"))
ps
```
and compute the Groebner basis
```{r}
gb <- groebner(ps)
gb
```

Compute the monomial basis, the multiplication matrices and their common eigenbasis
```{r}
ms <- monomial_basis(gb)
Ms <- lapply(1:3,function(k) multiplication_matrix(k,ms,gb))
Bs <- common_eigenbasis(Ms)
```

The roots of the system are
```{r}
rs <- roots(Ms,Bs)
rs
```

and the total absolute error at the roots is
```{r}
rowSums(abs(eval_polys(ps,rs)))
```


## Example 3

Consider the system of polynomials
$$
\begin{gather*}
x^3 - 2 x y \\
x^2 y - 2 y^2 + x
\end{gather*}
$$

Define the polynomials
```{r}
set_term_order(grevlex)
ps <- parse_polys(c("x^3 - 2*x*y",
                    "x^2*y-2*y^2+x"),
                  c("x","y"))
ps
```

Compute the Groebner basis
```{r}
gb <- groebner(ps)
gb
```

In this case, the multiplication matrices are deficient
```{r}
ms <- monomial_basis(gb)
Ms <- lapply(1:2,function(k) multiplication_matrix(k,ms,gb))
Ms
```

The common eigenbasis is of dimension 1
```{r}
Bs <- common_eigenbasis(Ms)
Bs
```

so there is only a single root
```{r}
rs <- roots(Ms,Bs)
rs
```

The total absolute error at the root is
```{r}
rowSums(abs(eval_polys(ps,rs)))
```


## Example 4

Consider the system of polynomials
$$
\begin{gather*}
x^3 - 5 x^2 + 8 x - 4 \\
x^2 y + 2 x^2 - 12 x - 4 y + 16 \\
xy + x^2 - 5 x - 2 y + 6 \\
y^2 + 2 x - 4 y + 4
\end{gather*}
$$


Define the polynomials
```{r}
set_term_order(grevlex)
ps <- parse_polys(c("x^3 - 5*x^2 + 8*x - 4",
                    "x^2*y + 2*x^2 - 12*x - 4*y + 16",
                    "x*y + x^2 - 5*x - 2*y + 6",
                    "y^2 + 2*x - 4*y + 4"),
                  c("x","y"))
ps
```

Compute the Groebner basis
```{r}
gb <- groebner(ps)
gb
```

The roots of the system are
```{r}
rs <- solve_polys(ps,gb)
rs
```

The total absolute error at the roots is
```{r}
rowSums(abs(eval_polys(ps,rs)))
```


## Example 5

Consider the system of polynomials
$$
\begin{gather*}
x^3 - 5 x^2 + 8 x - 4 \\
x^2 y + 2 x^2 - 12 x - 4 y + 16 \\
xy + x^2 - 5 x - 2 y + 6 \\
y^2 - x^2 + 4 x - 2 y - 3
\end{gather*}
$$

Define the polynomials
```{r}
set_term_order(grevlex)
ps <- parse_polys(c("x^3 - 5*x^2 + 8*x - 4",
                    "x^2*y + 2*x^2 - 12*x - 4*y + 16",
                    "x*y + x^2 - 5*x - 2*y + 6",
                    "y^2 - x^2 + 4*x - 2*y - 3"),
                  c("x","y"))
ps
```

Compute the Groebner basis
```{r}
gb <- groebner(ps)
gb
```

The roots of the system are
```{r}
rs <- solve_polys(ps,gb)
rs
```

The total absolute error at the roots is
```{r}
rowSums(abs(eval_polys(ps,rs)))
```
